Title:      Reproducible Research
=================================

I split this assignment into two parts.
This is the analysis of the data with the NA values omitted.
The program name is 'prog_01.Rmd' with corresponding 'prog_01.html' file.
=========================================================================

There is a second program that produces the analyis for data with the NAs
replaced by the average for that interval over the two months.
The program name is 'prog_02.Rmd' with corresponding 'prog_02.html' file.

==========================================================================
```{r, echo=TRUE}
    library(dplyr)
    library(knitr)
    library(ggplot2)
```
Loading and preprocessing the data. Show any code that is needed to.
Load the data (i.e. read.csv()).
Process/transform the data (if necessary) into a format suitable for your analysis.
===================================================================================

The date and time were reformated. An new datetime column was created.


```{r}
    opts_chunk$set(echo = TRUE, results = 'hold')
    act_01   <-    read.csv("activity.csv", header=TRUE)
    summary(act_01)
    head(act_01)
    NROW(act_01)

    print (sapply(act_01, typeof))

    act_02 <- filter(act_01)  %>% na.omit()
    NROW(act_02)

    act_03 <- mutate(act_01, interval, na.rm=TRUE, date, steps,
                     Date_01 = format.Date(date, '%Y-%m-%d'),
                     Time_01 = paste0(substr(sprintf('%04d', interval), 1,2),                                               ':',
                                      substr(sprintf('%04d', interval), 3, 4)),
                     Datetime_01 =    as.POSIXct(paste0(Date_01, Time_01)))

    summary(act_03)
    head(act_03)
    NROW(act_03)
    print (sapply(act_03, typeof))
    
    act_04 <- aggregate(steps ~ date, act_03, sum)
    
    colnames(act_04) <- c("Date","Steps")

    head(act_04)
    NROW(act_04)
```

What is the mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

Calculate the total number of steps taken per day.

Here is a breakdown of the number of steps per day.
===================================================
Each day has 24 x 12 five minute intervals = 288.
There are 61 days from 1st October to 30 Nomember.
NA values are excluded.
===================================================

```{r}
    kable(act_04)
```
Take the column "steps" from the "act_04" dataset and make a histogram of it.
=============================================================================

Comments:
=========

The histogram suggests a Gaussian distribution with a possible positive skew.
The intervals with zero steps (during sleep) may need to be ommitted so that the 
data reflects true activity during the day-time.

The mean and median are pracically identical.


```{r}
    paddy_01 <- ggplot(data=act_04, aes(act_04$Steps)) + 
                geom_histogram(breaks=seq(0, 21000, by = 1000), 
                 col="red", 
                 fill="green", 
                 alpha = 1) + 
  labs(title="Histogram for Steps") +
  labs(x="Steps", y="Count") + 
  xlim(c(0,21000)) + 
  ylim(c(0, 11))
    
    paddy_01 + theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=14, angle=45),
                     axis.text.y = element_text(face="bold", color="#993333", 
                           size=14, angle=45))

    round(mean(act_04$Steps) )
    median(act_04$Steps)
    
```
What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Comments:
=========

The data seems to make sense. People go to bed from 10pm onwards.
Most people sleep between 12 midnight and 5am.
People start getting up from 5am onwards and possibly walk to work.
The data suggests a sedentary group of workers. There is a spike about lunch time.
People leave work at different times or are involved in different activities in the evening.
========================================================================================
```{r}
    act_05 <- aggregate(steps ~ interval, act_03, sum)
    head(act_05)
    NROW(act_05)

    paddy_02 <- ggplot(act_05, aes(interval, steps)) + geom_line() +
                xlab("") + ylab("Average steps in 5 minute intervals")
    
    paddy_02
    
    act_06 <- filter(act_05, steps, interval, steps == max(steps))
    head(act_06)
    NROW(act_06)
```

The end of the world is nigh...........!!!!!!!!!!!!!!!!!!!